Tail-indices and scale parameters of returns series

### 2017-04-05 cv/20sec/template.tex
Add parallel computing using OpenMP to my skill list.
Wrote an abstract of the FX paper and revised my CV accordingly.

% Skill bar section, each skill must have a value between 0 an 6 (float)
\skills {{Python/3},{SQL/4},{GNU Scientific Library/5},{Matlab/5.2},{R/5.8},{C/5},{C++/5}}

@@

\section{Papers in Finalization}
\begin{itemize}
\item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
  {\it Generalized Disappointment Aversion and heavy-tail property of
    equity return distributions}

  {\small
    We study the heavy tail properties of real-world equity return
    distributions and, based on the framework of Generalized
    Disappointment Aversion, derive how an investor's preference varies
    with the tail distribution characteristics, namely the tail index and
    the scale parameter. We find, when the equity return is t-distributed,
    the GDA preference actually decreases with the degree of freedom of
    the distribution, i.e. the tail index; we interprete this as a
    consequence of the symmetry of the distribution. When the equity
    return is assumed to have a shifted Pareto distribution with
    independent parameters of the left and the right tails, we find the
    GDA preference increasing with both the lower tail index and the scale
    parameter. We attribute this to the risk avert nature of rational
    investors. By numerical methods, optimal asset allocation is also
    solved in both cases.
  }
\end{itemize}
%----------------------------------------------------------------------------------------
%	 AWARDS
%----------------------------------------------------------------------------------------

>>>>

% Skill bar section, each skill must have a value between 0 an 6 (float)
\skills {{Python/3},{SQL/4},{Matlab/5.2},{GNU Scientific Library/5},{R/5.8},{Parallel Computing OpenMP/4.5},{C/5},{C++/5}}

@@

\section{Papers in Finalization}
\begin{itemize}
\item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
  {\it Tail-indices and scale parameters of returns series}

  {\small
    %% We study the heavy tail properties of real-world equity return
    %% distributions and, based on the framework of Generalized
    %% Disappointment Aversion, derive how an investor's preference varies
    %% with the tail distribution characteristics, namely the tail index and
    %% the scale parameter. We find, when the equity return is t-distributed,
    %% the GDA preference actually decreases with the degree of freedom of
    %% the distribution, i.e. the tail index; we interprete this as a
    %% consequence of the symmetry of the distribution. When the equity
    %% return is assumed to have a shifted Pareto distribution with
    %% independent parameters of the left and the right tails, we find the
    %% GDA preference increasing with both the lower tail index and the scale
    %% parameter. We attribute this to the risk avert nature of rational
    %% investors. By numerical methods, optimal asset allocation is also
    %% solved in both cases.
    It is a stylized fact that equity return distributions are
    heavy-tailed. We discuss how an investor's preference of an equity
    varies with the parameters of the tail distribution of the return on
    equity. These parameters can be seen as characteristics of the
    market and provide a quantitative description of the status of the
    market.
    In addition, we also find optimal asset allocation schemes
    corresponding to different investor preferences as captured by the
    investors' utility functions. These results can be useful when
    considering a problem of asset allocation. Real-world stock return
    series, which have motivated our work, are analyzed and presented.
  }
\end{itemize}
%----------------------------------------------------------------------------------------
%	 AWARDS
%----------------------------------------------------------------------------------------

### 2017-04-05 papers/FX/draft.tex
Wrote an abstract of the FX paper and revised my CV accordingly.

\allowdisplaybreaks
\begin{document}
\title{Tail-indices and scale parameters of returns series}
\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei, }
\date{\today}

@@

\begin{abstract}
%We consider an investor with preferences that accord with Generalized
%Disappointment Aversion. Such an investor cares about downside risk and
%we assume he recognizes the heavy tail feature of asset return
%distributions. We argue that when a market is dominated by rational
%investors of this kind, the return distributions of equities that are
%actively traded in this market must have very similar tail indices. We
%also show, in contrast, the scale parameters of the return
%distributions may differ hugely from one another.
%On the other hand, whether or not all equities in a multivariate model
%have the same tail index is a dividing issue for multivariate GARCH
%models proposed in the literature. Therefore, it is important to analyze
%data of real equity returns and see how close to each other the
%tail indices actually are.
%In this work empirical results are also presented and they appear
%to support the conclusion that the tail indices are very similar,
%with respect to the confidence bands of estimation.
\end{abstract}

@@

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1_{\{x < q\}} \over (K' + x)^{\beta + 1}} dx 
  - b u(\delta v) F(q)
  % &&
  % - b u(\delta v) \left[
  %   1 - (1 - p) {
  %     K'^\beta
  %     \over
  %     (K' + q)^\beta
  %   }
  % \right]
  \label{eq:xxie1.0}
\end{eqnarray}
\begin{theorem}
  \label{thrm:I}
  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
  function $u(\cdot)$ is increasing and differentiable, then
  \[
  \frac{\pd G(F)}{\pd \alpha} > 0,
  \quad
  \frac{\pd G(F)}{\pd K} < 0
  \]
\end{theorem}
The proof of theorem \ref{thrm:I} is given in secton
\ref{sec:thrmI_proof}.
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of S\&P 500 companies. This is
illustrated in figure \ref{fig:sectors_parameters}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figure~\ref{fig:1}.

@@

\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_p.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
  \label{fig:sectors_parameters}
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
% Comparing this expression with \eqref{eq:lower_tail_expansion},
Then one can identity $p K^\alpha$ as the scale parameter. Figure
\ref{fig:scale_estimates} compare the estimates obtained
from Hill's scale estimator with those from the $p K^\alpha$ method.

>>>>

\allowdisplaybreaks
\begin{document}
\title{Tail-indices and scale parameters of returns series}
\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
\date{\today}

@@

\begin{abstract}
  It is a stylized fact that equity return distributions are
  heavy-tailed. We discuss how an investor's preference of an equity
  varies with the parameters of the tail distribution of the return on
  equity. These parameters can be seen as characteristics of the
  market and provide a quantitative description of the market's status.
  In addition, we also find optimal asset allocation schemes
  corresponding to different investor preferences as captured by the
  investors' utility functions. These results can be useful when
  considering a problem of asset allocation. Real-world stock return
  series, which have motivated our work, are analyzed and presented.

%We consider an investor with preferences that accord with Generalized
%Disappointment Aversion. Such an investor cares about downside risk and
%we assume he recognizes the heavy tail feature of asset return
%distributions. We argue that when a market is dominated by rational
%investors of this kind, the return distributions of equities that are
%actively traded in this market must have very similar tail indices. We
%also show, in contrast, the scale parameters of the return
%distributions may differ hugely from one another.
%On the other hand, whether or not all equities in a multivariate model
%have the same tail index is a dividing issue for multivariate GARCH
%models proposed in the literature. Therefore, it is important to analyze
%data of real equity returns and see how close to each other the
%tail indices actually are.
%In this work empirical results are also presented and they appear
%to support the conclusion that the tail indices are very similar,
%with respect to the confidence bands of estimation.
\end{abstract}

@@

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  +
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1_{\{x < q\}} \over (K' + x)^{\beta + 1}} dx 
  - b u(\delta v) F(q)
  % &&
  % - b u(\delta v) \left[
  %   1 - (1 - p) {
  %     K'^\beta
  %     \over
  %     (K' + q)^\beta
  %   }
  % \right]
  \label{eq:xxie1.0}
\end{eqnarray}
\begin{theorem}
  \label{thrm:I}
  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
  function $u(\cdot)$ is increasing and differentiable, then
  \[
  \frac{\pd G(F)}{\pd \alpha} > 0,
  \quad
  \frac{\pd G(F)}{\pd K} < 0
  \]
\end{theorem}
The proof of theorem \ref{thrm:I} is given in secton
\ref{sec:thrmI_proof}.
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of S\&P 500 companies. This is
illustrated in figure \ref{fig:sectors_parameters}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figure~\ref{fig:1}.

@@

\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_p.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_p.pdf}
  \end{minipage}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_alpha_K_ci.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_p.pdf}
  \end{minipage}
  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
    S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals.
  }
  \label{fig:sectors_parameters}
\end{figure}
To compare the maximum likelihood estimates with the results obtained
from Hill's scale estimates, one can re-write the lower-tail
distribution function: For all $x < 0$,
\begin{eqnarray*}
  \P(X < x)
  &=&
  p {K^\alpha \over (K - x)^\alpha} \\
  &=&
  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
  &=&
  p K^\alpha |x|^{-\alpha} \left(
  1 + {\alpha K \over x} + O(|x|^{-2})
  \right)
\end{eqnarray*}
% Comparing this expression with \eqref{eq:lower_tail_expansion},
Then one can identity $p K^\alpha$ as the scale parameter. Figure
\ref{fig:scale_estimates} compare the estimates obtained
from Hill's scale estimator with those from the $p K^\alpha$ method.
